---
title: "Homework3CarolineSeip"
author: "Caroline"
date: "September 27, 2018"
output: github_document
---

#Table of contents
- Introduction
- Loading packages
- Task 1: Get the maximum and minimum of GDP per capita for all continents
- Task 2: How is life expectancy changing over time on different continents?
- Task 3
- But I want to do more!

First let's load the dataset, tidyverse and knitr
```{r}
#Load gapminder dataset
library(gapminder)
#Load tidyverse to use dplyr and ggplot
library(tidyverse)
#Load knitr to use the kable function
library(knitr)
```

#Task 1: Get the maximum and minimum of GDP per capita for all continents

```{r}
#Use the gapminder dataset
a <- gapminder %>% 
  #Group by continent
  group_by(continent) %>%
  #Use mutate and summarize to create minimum and maximum variables
  mutate_each(funs(round(.,2)), gdpPercap) %>% 
  summarize(min=min(gdpPercap),
            max=max(gdpPercap))
#Now put this into a table using the 'kable' function
a %>% 
kable(col.names = c("Continent", "Minimum GDP per capita", "Maximum GDP per capita"))
```

Ok, now let's do a check to make sure that went right... let's look at the minimum GDP per capita for Oceania
```{r}
#Use the gapminder dataset
gapminder %>% 
  #Filter to only see Oceania data
  filter(continent == "Oceania") %>% 
  #Arrange by gdpPercap so that we can see the minimum first
  arrange(gdpPercap)
  
```

Yay it's right! We can see the GDP per capita for Australia in 1952 is 10040, which matches up with the summarize function we used to find the minimum GDP per capita in Oceania.

Let's do a visual check by making a box plot to see the minimum and maximum values of GDP per capita by continent:
```{r}
  #Use ggplot, specify your dataset and aesthetics(x,y)
ggplot(gapminder, aes(continent, gdpPercap)) +
  #Add a boxplot
  geom_boxplot(aes(fill=continent)) +
  #Label the axes
  xlab("Continent") +
  ylab("GDP per capita") +
  guides(fill=guide_legend(title= "Continent"))
```

#Task 2: How is life expectancy changing over time on different continents?


```{r}
#Use the gapminder dataset
long <- gapminder %>% 
  #Select the continent, year, and life expectancy variables
   select(continent, year, lifeExp) %>% 
  #Group by continent and year
  group_by(continent, year) %>% 
  #Find mean life expectancy for each continent, in each year
  summarize(life = mean(lifeExp))

wide <- spread(long, year, life)

colnames(wide) <- c("continent", "first", "last")
```

```{r}
gapminder %>%
  ggplot(aes(year, lifeExp)) +
  geom_smooth(se= FALSE, aes(colour= continent)) +
  xlab("Year") +
  ylab("Life expectancy (years)") +
  guides(fill=guide_legend(title= "Continent"))
```

