---
title: "Homework3CarolineSeip"
author: "Caroline"
date: "September 27, 2018"
output: github_document
---

#Table of contents
- Introduction
- Loading packages
- Task 1: Get the maximum and minimum of GDP per capita for all continents
- Task 2
- Task 3
- But I want to do more!

First let's load the dataset, tidyverse and knitr
```{r}
#Load gapminder dataset
library(gapminder)
#Load tidyverse to use dplyr and ggplot
library(tidyverse)
#Load knitr to use the kable function
library(knitr)
```

#Task 1: Get the maximum and minimum of GDP per capita for all continents

```{r}
a <- gapminder %>% 
  group_by(continent) %>%
  mutate_each(funs(round(.,2)), gdpPercap) %>% 
  summarize(min=min(gdpPercap),
            max=max(gdpPercap))
a %>% 
kable(col.names = c("Continent", "Minimum GDP per capita", "Maximum GDP per capita"))
```

Hmm let's do a check to make sure that went right... let's look at the minimum GDP per capita for Oceania
```{r}
gapminder %>% 
  #Filter to only see Oceania data
  filter(continent == "Oceania") %>% 
  #Arrange by gdpPercap so that we can see the minimum first
  arrange(gdpPercap)
  
```

Yay it's right!

Let's do a visual check by making a box plot to see the minimum and maximum values of GDP per capita by continent:
```{r}
  #Use ggplot, specify your dataset and aesthetics(x,y)
ggplot(gapminder, aes(continent, gdpPercap)) +
  #Add a boxplot
  geom_boxplot(aes(fill=continent)) +
  #Label the axes
  xlab("Continent") +
  ylab("GDP per capita") +
  guides(fill=guide_legend(title= "Continent"))
```

#Task 2: How is life expectancy changing over time on different continents?
```{r}
gapminder %>%
mean(lifeExp)
```
```{r}
gapminder %>% 
   select(continent, year, lifeExp) %>% 
  group_by(continent, year) %>% 
  summarize(life = mean(lifeExp)) %>% 
  kable()
```

```{r}
gapminder %>%
  ggplot(aes(year, lifeExp)) +
  geom_smooth(se= FALSE, aes(colour= continent)) +
  xlab("Year") +
  ylab("Life expectancy (years)") +
  guides(fill=guide_legend(title= "Continent"))
```

