---
title: "Homework3CarolineSeip"
author: "Caroline"
date: "September 27, 2018"
output: github_document
---

#Table of contents
-Introduction
-Loading packages
-Task 1: Get the maximum and minimum of GDP per capita for all continents
-Task 2
-Task 3
-But I want to do more!

First let's load the dataset, tidyverse and knitr
```{r}
#Load gapminder dataset
library(gapminder)
#Load tidyverse to use dplyr and ggplot
library(tidyverse)
#Load knitr to use the kable function
library(knitr)
```

#Task 1: Get the maximum and minimum of GDP per capita for all continents

```{r}
gapminder %>% 
  group_by(continent) %>%
  mutate_each(funs(round(.,2)), gdpPercap) %>% 
  summarize(min=min(gdpPercap),
            max=max(gdpPercap)) %>% 
  kable(col.names = c("Continent", "Minimum GDP per capita", "Maximum GDP per capita"))
```

Hmm let's do a check to make sure that went right... let's look at the minimum GDP per capita for Oceania
```{r}
gapminder %>% 
  #Filter to only see Oceania data
  filter(continent == "Oceania") %>% 
  #Arrange by gdpPercap so that we can see the minimum first
  arrange(gdpPercap)
  
```

```{r}
ggplot
```

